Title: Something snappy about the SLAM competition

### 2018-03-21 doc/naaclhlt2018-latex/paper.tex
add modeling section to paper

>>>>

After featurizing the training data, we trained gradient boosting decision
tree (GBDT) models to minimize log loss. GBDT works by
iteratively building regression trees, each of which seeks to minimize the
residual loss from prior trees. This allows it to capture non-linear effects
and high-order interactions among features. We used the LightGBM\footnote{http://lightgbm.readthedocs.io/} implementation
of GBDT \cite{ke2017lightgbm}.

For continuous-valued features, GBDT can split a leaf at any point, creating
different predicted values above and below that threshold. For categories that
are one-hot encoded, it can split a leaf on any of the category's features. This
means that for a category with thousands of values, potentially thousands of
tree splits would be needed to capture its relation to the target. Fortunately,
LightGBM implements an algorithm for partitioning the values of a categorical
feature into two groups based on their relevence to the current loss, and create
a single split to divide those groups \cite{fisher1958grouping}. Thus, as
alluded to above, high-cardinality features like token and user were encoded as
single columns and handled as categories by LightGBM.

We trained a model for each of the three language tracks of {\tt en\_es}, {\tt es\_en},
and {\tt fr\_en}, and also trained a model on the combined data from
all three tracks, adding an additional ``language'' feature. Following model
training, we averaged the predictions of each single-language model with that of
the all-language model to form our final predictions.

To tune model hyper-parameters and evaluate the usefulness of features, we first
trained the models on the ``train'' data set and evaluated them on the ``dev''
data set. Once the model structure was finalized, we trained on the combined
``train'' and ``dev'' data and produced predictions for the ``test'' data. The
LightGBM hyperparameters used for each model are listed in Table~\ref{lightgbm-params}.

\begin{table}[t!]
\begin{center}
\begin{tabular}{|l|llll|}
  % \hline   & \multicolumn{3}{l}{\bf Value} & \\
  \hline \bf Parameter & {\tt fr\_en} & {\tt en\_es} & {\tt es\_en} & {\tt all} \\ \hline
  num\_leaves & 256 & 512 & 512 & 1024 \\
  learning\_rate & .05 & .05 & .05 & .05 \\
  min\_data\_in\_leaf & 100 & 100 & 100 & 100 \\
  num\_boost\_rounds & 750 & 650 & 600 & 750 \\
  cat\_smooth & 200 & 200 & 200& 200 \\
  feature\_fraction & .7 & .7 & .7 & .7 \\
  max\_cat\_threshold & 32 & 32 & 32& 64 \\

\hline
\end{tabular}
\end{center}
\caption{\label{lightgbm-params} Parameters of final LightGBM models. See
  LightGBM documentation for more information; all other parameters were left at
their default values.}
\end{table}

\subsection{Performance}

The AUROC of our final predictions was $.8585$ on {\tt en\_es}, $.8350$ on {\tt
  es\_en}, and $.8540$ on {\tt fr\_en}. We did not attempt to optimize the model's
F1 score, but the model's F1 score could likely be improved (at the cost of
increased log loss) by finding the
rescaling of the ``dev'' predicted probabilities that maximized the F1 score at
the 0.5 threshold, and applying this rescaling to the ``test'' predicted probabilities.
